\documentclass[12pt]{article}
    \usepackage{style}
    \newcommand{\lang}{en}

    \usepackage{etoolbox}
    \apptocmd{\sloppy}{\hbadness 10000\relax}{}{} % Fixes under-full box in bibliography

    \usepackage{subfiles} % Load last

    \overfullrule=5pt
    \pgfplotsset{compat=newest}
    \pgfplotsset{scaled y ticks=false}
    
    % Paper specific commands
    \def\plotpath{../../../plots}
    \newcommand{\control}[1]{#1}

    % Title page
    \author{Andrea Titton\thanks{CeNDEF, Faculty of Economics and Business, University of Amsterdam
    \href{mailto:a.titton@uva.nl}{\texttt{a.titton@uva.nl}}}}
    \title{Regret and Climate Tipping Points\thanks{
      I thank my supervisors Florian Wagener and Cees Diks for the patient guidance on this paper. I also thank Rick van der Ploeg, Christoph Hambel, and Frank Venmans for the insightful discussions. Finally, I thank attendees of the EEA-ESESM conference, Rotterdam, 2024 for the constructive comments.
    }}

\begin{document}

\maketitle

\iffalse
\begin{abstract}
  Climate tipping points can lead to abrupt climate change and the costs associated with it.
\end{abstract}
\fi

\newpage
\subfile{sections/introduction/introduction.tex}

\section{Model}

This section introduces the climate model and the economy.

\subsection{Climate Model}

\subfile{sections/climate-model/climate-model.tex}

\subsection{Economy}

\subfile{sections/economy-model/economy-model.tex}


\section{Social Planner Problem and Regret}

\subfile{sections/social-planner/social-planner.tex}

\section{Results}

This section presents the optimal policies and the regret under three different scenarios. First, the case of an \textit{imminent} tipping point, $T^c = 1.5^{\circ}$. Second, the case of a \textit{remote} tipping point, $T^c = 2.5^{\circ}$. Third, the case of the literature \textit{benchmark} stochastic tipping model (as described in Appendix \ref{appendix:benchmark}).

\subfile{sections/results/optimal-abatement.tex}
\subfile{sections/results/regret.tex}


\newpage
\section{Conclusion}

This paper studies the role of tipping points in determining optimal emissions. Building on the calibration by \cite{hambel_optimal_2021}, I extend the climate dynamics to include a potential bifurcation induced by the loss in albedo due to the change in the area of ice caps, sea ice, and glaciers \cite{ashwin_tipping_2012,ashwin_extreme_2020}. I show that, in the presence of tipping points, optimal abatement is more ambitious in scope and timing. In fact, early abatement is crucial to avoid long periods of exposure to tipping risk.

The model presented here represents an early and simplified analysis that can be extended in various directions. First, more work is needed to analytically link the risk of tipping and the optimal abatement strategy, in order to quantify precisely the role of higher order climate dynamics in determining the social cost of carbon. Second, the underlying assumption of the social planner's optimisation problem is that she knows the climate dynamics and the role of the ice-albedo feedback. Such an assumption calls for extending the analysis to a situation in which the magnitude of the albedo loss is not known and rather can be estimated using early warning signals. Yet, in the face of uncertainty, the optimal abatement policy derived in this paper serves as a good rule against the possible worst case scenario.

\newpage
\bibliography{scc-tipping-points}

\newpage\appendix
\subfile{sections/appendix/climate.tex}
\subfile{sections/appendix/solution.tex}
\subfile{sections/appendix/calibration.tex}
\subfile{sections/appendix/benchmark.tex}

\end{document}